---
title: "week_01_homework"
output: html_notebook
---



```{r}
library(tidyverse)
library(readxl)
library(janitor)
library(stringr)
library(lubridate)
```

```{r}
books <- read.csv("data/books.csv")
glimpse(books)
view(books)
class(books)
```

```{r}

books %>% 
  # filter for high average_rating, <= 500 pages
  filter(average_rating >= 4.5, num_pages <= 500) %>% 
  #output = title, authors, no. of pages, no. of reviews
select(title, authors, num_pages, text_reviews_count)
```

.....
#### NOT WORKING!!

```{r}
books %>% 
  select(title %in% "Harry Potter")
```


```{r}
filter(books, title %in% c("The Lord of the Rings (The Lord of the Rings  #1-3)","In a Sunburned Country"))



dplyr::filter(books, title == "Bill Bryson's African Diary")
```
```{r}

dplyr::filter(books, title == "Bill Bryson's African Diary")
```
  
  
  
```{r}
books %>% 
  filter(authors == "J.K. Rowling")
```
  
  
  
```{r}
books %>% 
  filter(authors %in% "Jane Austen")
```
  
  
  
```{r}
#selecting books by searching for partial string
books %>% 
  filter(str_detect(authors, "Jane Austen")) %>% 
select(title, authors)
```
  
  
..........
# 1.  Looking at the correlation between number of pages and rating
  
```{r}
ggplot(books, aes(x = average_rating, y = num_pages)) +
  geom_col()
```


# Noticed that there are a number of ratings of 0 - strip those out

```{r}
true_ratings <- books %>% 
  filter(average_rating > 0) 
  ggplot(true_ratings, aes(x = average_rating, y = num_pages)) +
  geom_col()
```

  # Change y axis to log10
  
```{r}
 ggplot(true_ratings, aes(x = average_rating, y = num_pages)) +
  geom_col() +
  labs(scale_y_log10)

ggplot(true_ratings, aes(x = average_rating, y = num_pages)) +
geom_col() +
scale_y_continuous(trans='log10')


ggplot(true_ratings, aes(x = average_rating, y = num_pages, log10 = "y")) +
geom_col()
  
ggplot(true_ratings, aes(x = average_rating, y = num_pages, scale_y_log10())) +
geom_col()

```
  
  
  #adding some labels
  
```{r}
true_ratings <- books %>% 
  filter(average_rating > 0) 
  ggplot(true_ratings, aes(x = average_rating, y = num_pages)) +
  geom_col() +
    labs(
      x = "Average Book Rating",
      y = "Number of Pages",
      title = "Average Book Rating by Number of Pages")
```
  
  
  
.....


continent_avgs_2007 <- gapminder %>% 
  filter(year == 2007) %>% 
  group_by(continent) %>% 
  summarise(
    average_life_expectancy = mean(lifeExp),
    average_gdp_per_capita = mean(gdpPercap)
    )

continent_avgs_2007

```{r}
published_year_2007 <- books %>% 
  filter(str_detect(publication_date, "2007"))

published_year_2011 <- books %>% 
  filter(str_detect(publication_date, "2011"))

published_year_2007 %>% 
  filter(average_rating >= 4) +
  
```

 
....
# Lorna/Jamies code for Douglas Adams



library(lubridate)


```{r}
library(lubridate)

d_adams <- books %>%
  filter(authors == "Douglas Adams") %>%
  mutate(date = mdy(publication_date)) #change date format

d_adams %>%
  arrange(date)
## date is a character vector not dates

d_adams %>%
  ggplot() +
  aes(x = date, y = average_rating) +
  geom_line(col = "red") +
  geom_point(col = "red") +
  labs(
    x = "Publication Date",
    y = "Average Rating",
    title = "Douglas Adams Books",
    subtitle = "Publication Dates and Average Ratings"
  )
  
```




....# Ian/Jamies code to change the publication language to be able to filter out those is english language (includcing en_US, en_CA)


```{r}
books <- books %>%
  mutate(language_code_group = if_else(
    language_code %in% c("eng", "en-US", "en-GB", "en-CA"),
    "eng",
    language_code)
  )
```




  
  